---
title: "Final Project Dagitty Rough Draft"
author: "Daniel Lee"
date: "April 6, 2017"
output: pdf_document
---

##Causal Relationship between Salary and Whether a Person Leaves the Job or Not

Exposure variable $A$ is salary.
Outcome variable $Y$ is whether the person left the job or not.
Baseline covarates $W$ includes the following:

* work accident
* promotion
* last evaluation
* satisfaction
* department
* number of projects
* average monthly hours
* time spent at the company  

```{r , fig.height = 3, fig.width = 3, fig.align = "center"}

library(dagitty)

salary <- dagitty("dag { 
A <- U1 -> W <- U2 -> Y
A <- W -> Y
A <- U3 -> Y
A -> Y
}")
coordinates(salary) <- list(x=c(A=1,U3=2,Y=3,U1=1,U2=3,W=2),
                            y=c(A=0,U3=1,Y=0,U1=-1.25,U2=-1.25,W=-1))
exposures(salary) <- "A"
outcomes(salary) <- "Y"
plot(salary)

adjustmentSets(salary, effect = "direct")

```

To be able to identify the effect of salary on whether a person leaves the job or not, we make the following independence assumptions and condition on $W$:

$U_{A} \perp U_{Y}$, $U_{W} \perp U_{Y}$

```{r , fig.height = 3, fig.width = 3, fig.align = "center"}

salary_adjusted <- dagitty("dag { 
A <- U1 -> W
A <- W -> Y
A -> Y
U_Y -> Y
}")
coordinates(salary_adjusted) <- list(x=c(A=1,Y=3,U1=0.75,W=2, U_Y = 4),
                                     y=c(A=0,Y=0,U1=-0.25,W=-0.5, U_Y = 0))
plot(salary_adjusted)

```

\pagebreak

##Causal Relationship between Average Monthly Hours and Whether a Person Leaves the Job or Not

Exposure variable $A$ is average monthly hours a person works.
Outcome variable $Y$ is whether the person left the job or not.
Baseline covarates $W$ includes the following:

* work accident
* promotion
* last evaluation
* satisfaction
* department
* number of projects
* time spent at the company 
* salary  

```{r , fig.height = 3, fig.width = 3, fig.align = "center"}

average_monthly_hours <- dagitty("dag { 
A <- U1 -> W <- U2 -> Y
A <- W -> Y
A <- U3 -> Y
A -> Y
}")
coordinates(average_monthly_hours) <- list(x=c(A=1,U3=2,Y=3,U1=1,U2=3,W=2),
                            y=c(A=0,U3=1,Y=0,U1=-1.25,U2=-1.25,W=-1))
exposures(average_monthly_hours) <- "A"
outcomes(average_monthly_hours) <- "Y"
plot(average_monthly_hours)

adjustmentSets(average_monthly_hours, effect = "direct")

```

To be able to identify the effect of average monthly hours a person works on whether a person leaves the job or not, we make the following independence assumptions and condition on $W$:

$U_{A} \perp U_{Y}$, $U_{W} \perp U_{Y}$

```{r , fig.height = 3, fig.width = 3, fig.align = "center"}

average_monthly_hours_adjusted <- dagitty("dag { 
A <- U1 -> W
A <- W -> Y
A -> Y
U_Y -> Y
}")
coordinates(average_monthly_hours_adjusted) <- list(x=c(A=1,Y=3,U1=0.75,W=2, U_Y = 4),
                                     y=c(A=0,Y=0,U1=-0.25,W=-0.5, U_Y = 0))
plot(average_monthly_hours_adjusted)

```

\pagebreak

##Causal Relationship between Job Satisfaction Level and Whether a Person Leaves the Job or Not

Exposure variable $A$ is job satisfaction level.
Outcome variable $Y$ is whether the person left the job or not.
Baseline covarates $W$ includes the following:

* work accident
* promotion
* last evaluation
* salary
* department
* number of projects
* average monthly hours
* time spent at the company  

```{r , fig.height = 3, fig.width = 3, fig.align = "center"}

satisfaction_level <- dagitty("dag { 
A <- U1 -> W <- U2 -> Y
A <- W -> Y
A <- U3 -> Y
A -> Y
}")
coordinates(satisfaction_level) <- list(x=c(A=1,U3=2,Y=3,U1=1,U2=3,W=2),
                            y=c(A=0,U3=1,Y=0,U1=-1.25,U2=-1.25,W=-1))
exposures(satisfaction_level) <- "A"
outcomes(satisfaction_level) <- "Y"
plot(satisfaction_level)

adjustmentSets(satisfaction_level, effect = "direct")

```

To be able to identify the effect of job satisfaction level on whether a person leaves the job or not, we make the following independence assumptions and condition on $W$:

$U_{A} \perp U_{Y}$, $U_{W} \perp U_{Y}$

```{r , fig.height = 3, fig.width = 3, fig.align = "center"}

satisfaction_level_adjusted <- dagitty("dag { 
A <- U1 -> W
A <- W -> Y
A -> Y
U_Y -> Y
}")
coordinates(satisfaction_level_adjusted) <- list(x=c(A=1,Y=3,U1=0.75,W=2, U_Y = 4),
                                     y=c(A=0,Y=0,U1=-0.25,W=-0.5, U_Y = 0))
plot(satisfaction_level_adjusted)

```

\pagebreak

##Causal Relationship between Performance Evaluation and Whether a Person Leaves the Job or Not

Exposure variable $A$ is the last performance evaluation the person received at the job.
Outcome variable $Y$ is whether the person left the job or not.
Baseline covarates $W$ includes the following:

* work accident
* promotion
* salary
* satisfaction
* department
* number of projects
* average monthly hours
* time spent at the company  

```{r , fig.height = 3, fig.width = 3, fig.align = "center"}

last_evaluation <- dagitty("dag { 
A <- U1 -> W <- U2 -> Y
A <- W -> Y
A <- U3 -> Y
A -> Y
}")
coordinates(last_evaluation) <- list(x=c(A=1,U3=2,Y=3,U1=1,U2=3,W=2),
                            y=c(A=0,U3=1,Y=0,U1=-1.25,U2=-1.25,W=-1))
exposures(last_evaluation) <- "A"
outcomes(last_evaluation) <- "Y"
plot(last_evaluation)

adjustmentSets(last_evaluation, effect = "direct")

```

To be able to identify the effect of last evaluation on whether a person leaves the job or not, we make the following independence assumptions and condition on $W$:

$U_{A} \perp U_{Y}$, $U_{W} \perp U_{Y}$

```{r , fig.height = 3, fig.width = 3, fig.align = "center"}

last_evaluation_adjusted <- dagitty("dag { 
A <- U1 -> W
A <- W -> Y
A -> Y
U_Y -> Y
}")
coordinates(last_evaluation_adjusted) <- list(x=c(A=1,Y=3,U1=0.75,W=2, U_Y = 4),
                                     y=c(A=0,Y=0,U1=-0.25,W=-0.5, U_Y = 0))
plot(last_evaluation_adjusted)

```